ML Estimation of Matrix--$t$ Distributions with Expectation Maximization

### 2017-02-02 paper.tex
Begin working on sect3
Slightly more detailed wording.

\section{EM formulae for matrix--$t$ distribution}\label{sec:EM}

Given a random variable $X$ with probability density (or mass) function $p_{\theta}(\cdot)$. Assume the density (or mass) function is a marginalization over some unobserved random state $Z$, i.e., $p_{\theta}(x)=\mathbb{E}_{\theta}\left[p_{\theta}(x|Z)\right]$. With observations $\left\{x_i\right\}_{i\in I}$ the classical EM method for ML estimation consists of two steps:

\begin{itemize}
\item At the $k$-th iteration, compute the conditional expectations with respect to $Z$ for each observation $x_i$
\begin{equation}
Q_i(\theta,\theta_k)=\mathbb{E}_{\theta_k}\left[\log p_{\theta}(x_i,Z)|x_i\right].
\end{equation}

\item Update $\theta$ by maximizing $Q(\theta,\theta_k):=\sum_{i\in I}Q_i(\theta,\theta_k)$ with respect to $\theta$, i.e.,
\begin{equation}
\theta_{k+1} = \textrm{argmax}_{\theta}\,Q(\theta,\theta_k).
\end{equation}
\end{itemize}

@@

Given observed matrix data $\left\{T_j\right\}_{j\in J}$ for each observation $T_j$ the E-step of the matrix--$t$ distribution entails computing the following integral at step $k$
\begin{equation}
Q_j(\theta,\theta_k)=f(T_j;\theta_k)^{-1}\int_{\mathbf{S}_{++}^p}f(T_{j},S;\theta_k)\log f\left(T_{j},S;\theta\right)\,dS,\label{eq:expdef}
\end{equation}
where $\theta=\left\{M,\Sigma,\Omega,\nu\right\}$, $\theta_k=\left\{M_k,\Sigma_k,\Omega_k,\nu_k\right\}$, and $f(T,S;\theta)$ is the full likelihood of the distribution with the hidden scale parameter $S$.

>>>>

\section{EM formulae for the matrix--$t$ distribution}\label{sec:EM}

We begin by explaining how the EM method can be used to determine the
maximum likelihood parameters for some model from a set of
observations: Consider a random variable $X$ with probability density
(or mass) function $p_{\theta}(\cdot)$. The EM method applies if the
density is a marginalization over some unobserved random state $Z$,
i.e., $p_{\theta}(x)=\mathbb{E}_{\theta}\left[p_{\theta}(x|Z)\right]$.
Given a set of observations $\left\{x_j\right\}_{j\in J}$, the
classical EM method for ML estimation consists of two steps:
%
\begin{itemize}
\item At the $k$-th iteration, compute the conditional expectations
	with respect to $Z$ for each observation $x_j$ \begin{equation}
Q_i(\theta,\theta_k)=\mathbb{E}_{\theta_k}\left[\log p_{\theta}(x_i,Z)|x_i\right].
\end{equation}

\item Update $\theta$ by maximizing $Q(\theta,\theta_k):=\sum_{j\in J}Q_j(\theta,\theta_k)$ with respect to $\theta$, i.e.,
\begin{equation}
\theta_{k+1} = \textrm{argmax}_{\theta}\,Q(\theta,\theta_k).
\end{equation}
\end{itemize}
The algorithm is initialized by providing some initial guess for the
parameters of interest, $\theta_0$.

We now show how the EM method is applied to matrix--$t$ distributions.

@@

Given observed matrix data $\left\{T_j\right\}_{j\in J}$ (with a total
of $|J|$ observed matrix instances), for each observation $T_j$ the
conditional expectation at step $k$ is
%
\begin{equation}
Q_j(\theta,\theta_k)=f(T_j;\theta_k)^{-1}\int_{\mathbf{S}_{++}^p}%
f(T_{j},S;\theta_k)\log f\left(T_{j},S;\theta\right)\,dS%
,\label{eq:expdef}
\end{equation}
where $\theta=\left\{M,\Sigma,\Omega,\nu\right\}$,
$\theta_k=\left\{M_k,\Sigma_k,\Omega_k,\nu_k\right\}$, and
$f(T,S;\theta)$ is the full likelihood of the distribution with the
hidden scale parameter $S$.
